Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-09 paper/paper.tex
interm2

\subsection{The right and left TPJ: prediction error signaling and world semantics}
The DMN emerges with its midline structures early in human development
\citep{doria2010}, while
the right and left TPJs may become fully integrated into this major brain
network only after birth.
The TPJs are known to exhibit hemispheric differences
based on microanatomical properties and cortical gyrification patterns
\citep{seghier2013angular}.
Globally, neuroscientific investigations on hemispheric functional specialization
have highlighted the right cerebral hemisphere as dominant for attentional functions
and the left side for semantic functions
\citep{seghier2013angular, bzdok2013tpj, bzdok2016left,
stephan2007mechanisms}.

The TPJ in the right hemisphere (RTPJ) denotes a broad functional zone with
varying anatomical nomenclature \citep{marstpj2011, seghier2010, seghier2013angular}
that has been shown to be closely related to
multi-sensory prediction and prediction error signaling
\citep{downar2000multimodal, vetter2011, shulman2010, shulman2007}.
This DMN region is probably central for
action initiation during goal-directed psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Its involvement was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005}, and
detection of environmental changes across
visual, auditory, or tactile stimulation
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and predicted outcomes,
without necessarily relating these neural processes to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
Additionally, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates functional control of brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
Indeed, neurological patients with RTPJ damage have particular difficulties
with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events, the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
Viewed from a RL account,
the RTPJ might reflect an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with unexpected environmental stimuli and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

The left TPJ of the DMN (LTPJ),
in turn, may have a functional relationship to Wernicke's area
involved in semantic processes \citep{blumenfeld_book} and has
been described as ``a temporoparietal transmodal gateway for language''
by some investigators \citep{mesulam_book}.
Neurological patients with damage in Wernicke's area
have a major impairment of language comprehension
when listening to others or reading a book.
Patient speech
preserves natural rhythm and normal syntax, yet the
voiced sentences lack meaning (i.e., aphasia).
Abstracting from speech interpretations in linguistics
and neuropsychology,
the LTPJ appears to mediate access to and binding of world knowledge,
such as required during action considerations
\citep{binder2011neurobiology, seghier2013angular}.
Consistent with this view,
LTPJ damage in humans also entailed problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing explicit language content
\citep{varney1987locus}.
% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Inner speech also hinges on knowledge recall
about the physical and social world.
Indeed,
the internal production of
verbalized thought (``language of the mind'') was closely related to the LTPJ
in a pattern analysis of brain volume
\citep{geva2011neural}.
Further,
episodic memory recall and mental imagery to forecast future events strongly draw on
re-assembling world knowledge.
Isolated building blocks of world structure get rebuilt
in internally constructed mental scenarios that
guide present action choice,
weigh hypothetical possibilities, and forecast event outcomes.
%
As a candidate component of a RL agent,
neural processes in the LTPJ may
contribute to the automated predictions of the environment
by incorporating experience-derived building blocks of world regularities
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time-scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation for Monte Carlo sampling of events and outcomes.

\begin{figure}[!h]
  \centering
  \includegraphics[width=1.\linewidth]{POMDP_Fig.pdf}
  \caption{\textbf{Illustration of Partially Observable Markov Decision Process (POMDP).} Given the current state of the environment,
    the agent takes an action by following the policy matrix as updated by the Bellman equation. The agent receives a triggered reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
  \label{fig:rl}
\end{figure}

@@

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
are applicable to something as complex as human behavior. This class of
reinforcement-learning models has had numerous successes in diverse domains. For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics
\citep{ng2004,abbeel2004} and in challenging multistep strategy games
~\citep{mnih2015,silver2016mastering,pritzel2017neural}. An even recent work~\citep{robinowitz2018}
develops an MDP-based reasoning about future behavior
of other agents (the \textit{ToM-Net}).
The idea is to use meta-learning (learning to learn) to build strong priors
about the behavior of of a population of other agents.

@@

\paragraph*{Reinforcement learning in the brain.}
RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
Indeed,
it has been proposed \citep{gershman2015computational} that a core property of human intelligence is the improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments, a situation perfectly captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
can be a mapping between algorithmic aspects
underlying model-free and model-based RL learning
and neurobiological aspects underlying brain function.
The neurotransmitter dopamine could serve
as a 'teaching signal' to better estimate value associations
and action policies by modulating
synaptic plasticity in the reward-processing circuitry, including the NAc.
In contrast, model-based RL would start off with some mechanistic assumptions about the dynamics of the world.
These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
reward contingencies, etc.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
An agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
offering relevant resources.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}
% Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
Such knowledge can be partly extracted from the environment: the agent infers a model of the world while learning to take optimal decisions based on the current representation of the environment. These methods learn what the effect is going to be of taking a particular action in a particular state. The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by a \textit{policy}, which maps states of the world
to probability distributions over candidate actions.
Starting at time $t=0$,
following a policy $\pi$ generates a trajectory of action choices:
\begin{eqnarray*}
  \begin{split}
    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect
      reward }R_0 = r(s_0, a_0, s_1)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect
      reward }R_1 = r(s_1, a_1,s_2)\\
    \vdots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and
      collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
    \vdots
  \end{split}
\end{eqnarray*}
We assume time-invariance
in that we expect the dynamics of the process
to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
Since an action executed in the present moment might have repercussions in the
far future, it turns out that the quantity to optimize is not the instantaneous
rewards $r(s, a)$, but a \textit{cumulative reward} estimate which takes into
account expected reward from action choices in the future. A common approach to
modeling this gathered outcome is the time-discounted cumulative reward% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable
  \footnote{Random as it depends both on the environment's dynamics and the
  policy $\pi$ being executed. The exponential delay discounting function used
  here refers to the usual formulation in the field of reinforcement learning,
  although psychological experiments may also reveal other discounting regimes
  \citep{green2004discounting}.}
measures the cumulative reward of
following an action policy $\pi$.
%
Note that value buffering may be realized in the vmPFC.
This DMN region has direct connections to the NAc, known to be involved in
reward evaluation~\citep{carlezon2009}.

@@

\paragraph*{A note on implicit and explicit memory.}
While Markov processes are usually memoryless, it is mathematically feasible to
incorporate the previous states of such model into the current state. This
extension may partially account for implicit memory at the behavioral level, but
may not explain the underlying neurobiological implementation or accommodate
explicit memory. Implicit memory-based processing arises in our MDP account of
DMN function in several different forms: successive updates of a) the action
policy and the value function, both being products of the past, as well as b)
the deep non-linear relationships within the hierarchical connections of
biological neural networks. The brain's adaptive synaptic connections can be
thought as a deep neural-network architecture affording an implicit form of
information compression of life experience. Such memory traces are stored in the
neural machinery and can be implicitly retrieved as a form of knowledge during
simulation of action rather than accessed as a stored explicit representation
\citep{pezzulo2011grounding}. c) Certain neural processes in the hippocampus can
be seen as some type of Monte Carlo sampling for memory recall, which can also
be a basis for probabilistic simulations across time-scales
\citep{schacter2007remembering,Axelrod2017}.

@@

In comparison,
both the semantic hypothesis and the present formal account based on MDPs
expose mechanisms of how action considerations could be explored.
In both accounts,
there is also little reason to assume that contemplating alternative realities of various
levels of complexity, abstraction, time-scale, and purpose
rely on mechanisms that are qualitatively different. This interpretation concurs with
DMN activity increases across time, space, and content domains
demonstrated in many brain-imaging studies
\citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
Further, the semantic hypothesis
and MDP account offer explanations why HC damage does
not only impair recalling past events, but also imagining hypothetical and future
scenarios \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-enabled, internally
generated information for probabilistic representation of action outcomes,
MDPs render explicit the grounds on which an action is
eventually chosen, namely, the estimated cumulative reward.
In contrast to many versions of the semantic hypothesis,
the MDPs naturally integrate the egocentric view
(more related to current action, state, and reward) and the
world view (more related to past and future actions, states, and rewards)
on the world in a same optimization problem.
Finally,
the semantic account of DMN function does not provide
suffcient explanation of \textit{how}
explicit world knowledge and logical analogies thereof
lead to foresight of future actions and states.
The semantic hypothesis does also not fully explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Our MDP account provides an algorithmic explanation in that
stochasticity of the parameter space exploration
by Monte Carlo approximation achieves better fine-tuning of the
action policies and inference of expected reward outcomes.
That is, the purposeful stochasticity of policy and value updates
in MDPs provides a candidate explanation for why humans may
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, mental scene construction according to the semantic
account is lacking an explicit time and incentive structure,
both of which are integral parts of the MDP interpretation of DMN function.

@@

A major hurdle in guessing DMN function from cognitive brain-imaging studies has been its similar neural engagement in different time-scales: thinking about the past (e.g., autobiographical memory retrieval), thinking about hypothetical presents (e.g., daytime mind-wandering), and
anticipating scenarios yet to come (e.g., delay discounting).
The MDP account of DMN activity offers a natural integration of a-priori diverging neural processes into a common framework.
%
It is an important advantage of the proposed artificial intelligence perspective
on DMN biology that it is practically computable and
readily motivates neuroscientific hypotheses that can be put to the test in future research.
%
Future neuroscience experiments on the DMN should operationalize
the set of action, value, and state variables governing
the behavior of intelligent RL agents.
%
At the least, we propose an alternative vocabulary to
describe, contextualize, and interpret experimental findings in neuroscience studies
on higher-level cognition.
%
Ultimately,
neural processes in the DMN may realize a brain-wide information integration
ranging from real experience over purposeful dreams to predicted futures
to continuously refine the organism's intervention on the world.

>>>>

\subsection{The right and left TPJ: prediction error signaling and world semantics}
The DMN emerges with its midline structures early in human development
\citep{doria2010}, while
the right and left TPJs may become fully functionally integrated into this major brain
network only after birth.
The TPJs are known to exhibit hemispheric differences
based on microanatomical properties and cortical gyrification patterns
\citep{seghier2013angular}.
In general, neuroscientific investigations on hemispheric functional specialization
have highlighted the right cerebral hemisphere as more dominant for attentional functions
and the left side more for semantic functions
\citep{seghier2013angular, bzdok2013tpj, bzdok2016left,
stephan2007mechanisms}.

The TPJ in the right hemisphere (RTPJ) denotes a broad functional zone with
varying anatomical nomenclature \citep{marstpj2011, seghier2010, seghier2013angular}
that has been shown to be closely related to
multi-sensory prediction and prediction error signaling
\citep{downar2000multimodal, vetter2011, shulman2010, shulman2007}.
This DMN region is probably central for
action initiation during goal-directed psychological tasks and for
sensorimotor behavior by integrating multi-sensory attention
\citep{corbetta2002control}.
Its involvement was repeatedly reported in
multi-step action execution \citep{hartmann2005takes},
visuo-proprioceptive conflict \citep{Balslev2005},
spatial reorienting \citep{corbetta2000},
and
detection of environmental changes across
visual, auditory, or tactile stimulation
\citep{downar2000multimodal}.
Direct electrical stimulation of the human
RTPJ during neurosurgery was associated with altered perception
and stimulus awareness \citep{blanke2002neuropsychology}.
%
It was argued that the RTPJ encodes actions and predicted outcomes,
without necessarily relating these neural processes to value estimation
\citep{liljeholm2013neural, hamilton2008action,
jakobs2009effects}.
More specifically, neural activity in the RTPJ has been proposed to reflect
stimulus-driven attentional reallocation to
self-relevant and unexpected sources of information
as a ‘‘circuit breaker’’ that recalibrates functional control of brain networks
\citep{bzdok2013tpj, corbettashul2008}.
% (Bzdok et al., 2013; Corbetta et al., 2008).
% Indeed, neurological patients with RTPJ damage have particular difficulties
% with multi-step actions \citep{hartmann2005takes}.
In the face of large discrepancies between actual and previously predicted
environmental events, the RTPJ acts as a potential switch between
externally-oriented mind sets focussed on the
sensory environment and internally-oriented mind sets focussed
on mental scene construction.
For instance, temporally induced RTPJ damage in humans diminished the
impact of predicted intentions of other individuals
\citep{young2010disruption},
a capacity believed to be enabled by the DMN.
Viewed from a RL account,
the RTPJ might reflect an important relay that shifts away
from the ‘‘internally directed’’ baseline processes
to, instead, deal with unexpected environmental stimuli and events.
% (TODO: slow endogeneous changes in state transitions by the policy and value matrices)

The left TPJ of the DMN (LTPJ),
in turn, may have a functional relationship to Wernicke's area
involved in semantic processes \citep{blumenfeld_book} and has
been described as ``a temporoparietal transmodal gateway for language''
by some investigators \citep{mesulam_book}.
Neurological patients with damage in this region
have a major impairment of language comprehension
when listening to others or reading a book.
Patient speech
preserves natural rhythm and normal syntax, yet the
voiced sentences lack meaning (i.e., aphasia).
Abstracting from speech interpretations in linguistics
and neuropsychology,
the LTPJ appears to mediate access to and binding of world knowledge,
such as required during action considerations
\citep{binder2011neurobiology, seghier2013angular}.
Consistent with this view,
LTPJ damage in humans also entailed problems in recognizing
others' pantomimed action towards objects
without obvious relation to processing explicit language content
\citep{varney1987locus}.
% (Varney and Damasio 1987; Rothi et al. 1991; Buxbaum et al. 2005).
%
Inner speech also hinges on knowledge recall
about the physical and social world.
Indeed,
the internal production of
verbalized thought (``language of the mind'') was closely related to the LTPJ
in a pattern analysis of brain volume
\citep{geva2011neural}.
Further,
episodic memory recall and mental imagery to forecast future events strongly draw on
re-assembling world knowledge.
Isolated building blocks of world structure get rebuilt
in internally constructed mental scenarios that
guide present action choice,
weigh hypothetical possibilities, and forecast event outcomes.
%
As a candidate component of a RL agent,
neural processes in the LTPJ may
contribute to the automated predictions of the environment
by incorporating experience-derived building blocks of world regularities
into ongoing action, planning, and problem solving.
% world -> policy matrix driven computation
% self -> value matrix driven computation

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation for Monte Carlo sampling of events and outcomes.

\begin{figure}[!h]
  \centering
  \includegraphics[width=1.\linewidth]{POMDP_Fig.pdf}
  \caption{\textbf{Illustration of a Partially Observable Markov Decision Process (POMDP).} Given the current state of the environment,
    the agent takes an action by following the policy matrix, which is iteratively updated by the Bellman equation. The agent receives a triggered reward and observes the next state. The process goes on until interrupted or a goal state is reached.}
  \label{fig:rl}
\end{figure}

@@

\paragraph*{Why Markov Decision Processes?} One may wonder whether MDP models
are applicable to something as complex as human behavior. This class of
reinforcement-learning models has had numerous successes in diverse domains. For instance, financial trading is largely a manifestation of strategic decision-making of interacting human agents. According to how the market responds, the agent incurs gain or loss as environmental feedback of the executed financial actions. Recent research on automatizing market exchanges by algorithmic trading has effectively deployed MDPs as a framework for modeling these elaborate behavioral dynamics  ~\citep{brazdil2017,yang15,yang14,yang12,dempster2006,hult2010, abergel2017}.
%
MDPs have also been effective as a behavioral model in robotics
\citep{ng2004,abbeel2004} and in challenging multistep strategy games
~\citep{mnih2015,silver2016mastering,pritzel2017neural}. 
More recent work has
developped an MDP-based way of reasoning about future behavior
of other agents \citep{robinowitz2018}.
The idea is to use meta-learning (learning to learn) to build strong priors
about the behavior of of a population of other agents.

@@

\paragraph*{Reinforcement learning in the brain.}
RL can be plausibly realized in the human brain \citep{doherty2015structure, daw2014algorithmic}.
Indeed,
previous authors have proposed
\citep{gershman2015computational} that a core property of human intelligence is the improvement
of expected utility outcomes as a strategy for action choice in uncertain
environments, a situation perfectly captured by the formalism of MDPs.
It has also long been proposed \citep{dayan2008decision} that there
can be a mapping between algorithmic aspects
underlying model-free and model-based RL learning
and neurobiological aspects underlying brain function.
The neurotransmitter dopamine could serve
as a 'teaching signal' to better estimate value associations
and action policies by modulating
synaptic plasticity in the reward-processing circuitry, including the NAc.
In contrast, model-based RL would start off with some mechanistic assumptions about the dynamics of the world.
These can be assumptions about the physical laws governing the agent's environment or constraints on the state space, transition probabilities between states,
reward contingencies, etc.
% For instance, if a rat in a maze knows
% that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
An agent might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a location
offering relevant resources.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}
% Model-based assumes knowledge of the environment in terms of the transition probabilities between states, the reward contingencies for state-action pairs, or any form of representation of the environment.
Such knowledge can be partly extracted from the environment: the agent infers a model of the world while learning to take optimal decisions based on the current representation of the environment. These methods learn what the effect is going to be of taking a particular action in a particular state. The result is an estimate of the underlying MDP which can then be either solved exactly or approximately, depending on the setting and what is feasible.

@@

\subsubsection{Accumulated rewards and policies}
The behavior of the agent is governed by a \textit{policy}, which maps states of the world
to probability distributions over candidate actions.
Starting at time $t=0$,
following a policy $\pi$ generates a trajectory of action choices:
\begin{eqnarray*}
  \begin{split}
    \textbf{choose action: }a_0 &\sim \pi(a|s_0)\\
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect
      reward }R_0 = r(s_0, a_0, s_1)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect
      reward }R_1 = r(s_1, a_1,s_2)\\
    \vdots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and
      collect reward }R_{t} = r(s_{t}, a_{t},s_{t+1})\\
    \vdots
  \end{split}
\end{eqnarray*}
We assume time invariance
in that we expect the dynamics of the process
to be equivalent over sufficiently long time windows of equal length (i.e., stationarity).
Since an action executed in the present moment might have repercussions in the
far future, it turns out that the quantity to optimize is not the instantaneous
rewards $r(s, a)$, but a \textit{cumulative reward} estimate which takes into
account expected reward from action choices in the future. A common approach to
modeling this gathered outcome is the time-discounted cumulative reward% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable
  \footnote{Random as it depends both on the environment's dynamics and the
  policy $\pi$ being executed. The exponential delay discounting function used
  here refers to the usual formulation in the field of reinforcement learning,
  although psychological experiments may also reveal other discounting regimes
  \citep{green2004discounting}.}
measures the cumulative reward of
following an action policy $\pi$.
%
Note that value buffering may be realized in the vmPFC by virtue of
this region's connections to the NAc of the reward sytem
\citep{carlezon2009}.

@@

\paragraph*{A note on implicit and explicit memory.}
While Markov processes are usually memoryless, it is mathematically feasible to
incorporate the previous states of such model into the current state. This
extension may partially account for implicit memory at the behavioral level, but
may not explain the underlying neurobiological implementation or accommodate
explicit memory. Implicit memory-based processing arises in our MDP account of
DMN function in several different forms: successive updates of a) the action
policy and the value function, both being products of the past, as well as b)
the deep non-linear relationships within the hierarchical connections of
biological neural networks. The brain's adaptive synaptic connections can be
thought as a deep neural-network architecture affording an implicit form of
information compression of life experience. Such memory traces are stored in the
neural machinery and can be implicitly retrieved as a form of knowledge during
simulation of action rather than accessed as a stored explicit representation
\citep{pezzulo2011grounding}. c) Certain neural processes in the hippocampus can
be seen as some type of Monte Carlo sampling for memory recall, which can also
be a basis for probabilistic simulations across time scales
\citep{schacter2007remembering,Axelrod2017}.

@@

In comparison,
both the semantic hypothesis and the present formal account based on MDPs
expose mechanisms of how action considerations could be explored.
In both accounts,
there is also little reason to assume that contemplating alternative realities of various
levels of complexity, abstraction, time scale, and purpose
rely on mechanisms that are qualitatively different. This interpretation concurs with
DMN activity increases across time, space, and content domains
demonstrated in many brain-imaging studies
\citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
Further, the semantic hypothesis
and MDP account offer explanations why HC damage does
not only impair recalling past events, but also imagining hypothetical and future
scenarios \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-enabled, internally
generated information for probabilistic representation of action outcomes,
MDPs render explicit the grounds on which an action is
eventually chosen, namely, the estimated cumulative reward.
In contrast to many versions of the semantic hypothesis,
the MDPs naturally integrate the egocentric view
(more related to current action, state, and reward) and the
world view (more related to past and future actions, states, and rewards)
on the world in a same optimization problem.
Finally,
the semantic account of DMN function does not provide
suffcient explanation of \textit{how}
explicit world knowledge and logical analogies thereof
lead to foresight of future actions and states.
The semantic hypothesis does also not fully explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Our MDP account provides an algorithmic explanation in that
stochasticity of the parameter space exploration
by Monte Carlo approximation achieves better fine-tuning of the
action policies and inference of expected reward outcomes.
That is, the purposeful stochasticity of policy and value updates
in MDPs provides a candidate explanation for why humans may
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, mental scene construction according to the semantic
account is lacking an explicit time and incentive structure,
both of which are integral parts of the MDP interpretation of DMN function.

@@

A major hurdle in guessing DMN function from cognitive brain-imaging studies has been its similar neural engagement in different time scales: thinking about the past (e.g., autobiographical memory retrieval), thinking about hypothetical presents (e.g., daytime mind-wandering), and
anticipating scenarios yet to come (e.g., delay discounting).
The MDP account of DMN activity offers a natural integration of a-priori diverging neural processes into a common framework.
%
It is an important advantage of the proposed artificial intelligence perspective
on DMN biology that it is practically computable and
readily motivates neuroscientific hypotheses that can be put to the test in future research.
%
Future neuroscience experiments on the DMN should operationalize
the set of action, value, and state variables governing
the behavior of intelligent RL agents.
%
At the least, we propose an alternative vocabulary to
describe, contextualize, and interpret experimental findings in neuroscience studies
on higher-level cognition.
%
Ultimately,
neural processes in the DMN may realize a brain-wide information integration
ranging from real experience over purposeful dreams to predicted futures
to continuously refine the organism's intervention on the world.
